---
title: "Quick Start"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{semfindr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Purpose

The `semfindr` package has functions for doing structural
equation modeling (SEM)) diagnostics, such as identifying
influential cases and computing various diagnostic measures.

# Workflow

Many of the measures for a case is computed by refitting
a model with this case removed. Instead of doing this once
for each measure, `semfindr` adopts this workflow:

- Decide cases to examine. All cases by default.

- For each selected case, remove it and refit the model.

- Store the results.

Functions for deletion statistics will compute them from
this list. Users can do as many diagnostic analysis as they
want without repeating the time consuming refitting step.

The following sections illustrates how to use some of the
major functions.

# Fitting the Target Model

The sample dataset is `semfindr::pa_dat2`, with variables
`iv1`, `iv2`, `m1`, and `dv`, and 100 cases. For
conveienice, we assign `pa_dat` to a new symbol, `dat`.

```{r}
library(semfindr)
dat <- pa_dat
head(dat)
```

Assume that the target model under examination is a path
model with two predictors, one mediator, and one outcome
variable:

```{r}
mod <-
"
m1 ~ iv1 + iv2
dv ~ m1
"
```

We fit the model by `lavaan::sem`:

```{r}
library(lavaan)
fit <- sem(mod, dat)
```

# Rerun *n* Times

We refit the model 100 times, each time with one case
removed:

```{r echo = FALSE, include = FALSE}
if (file.exists("semfindr_fit_rerun.RDS")) {
    fit_rerun <- readRDS("semfindr_fit_rerun.RDS")
  } else {
    fit_rerun <- lavaan_rerun(fit)
    saveRDS(fit_rerun, "semfindr_fit_rerun.RDS")
  }
```

```{r eval = FALSE}
fit_rerun <- lavaan_rerun(fit)
```

This example takes about 4 to 8 seconds. For larger samples
or more complicated models, `semfindr::lavaan_rerun` supports
parallel processing by setting `parallel` to `TRUE`.

`semfindr::lavaan_rerun` also supports selecting cases based
on Mahalanobis distance on all variables in the model or on
implied scores of outcome variables. See the help
page of `semfindr::lavaan_rerun` for the detail.

If this process is slow, we can save the results by 
`base::saveRDS` such that we can load it for analysis later.

# Diagnostic Functions

## Standardized Changes in Estimates

One intuitive measures is to compute the changes in parameter
estimates if a case is deleted, with the changes standardized
by their standard errors.

```{r}
fit_est_change <- est_change(fit_rerun)
round(head(fit_est_change), 3)
```

`semfindr::est_change` will compute the generalized Cook's
distance, `gcd` using the parameters examined. `gcd` is
analogous to Cook's distance in multiple regression,
measures the "overall" change in the parameters if a case
is deleted.

Let's examine the top five cases with largest `gcd`:

```{r}
i <- order(fit_est_change[, "gcd"],
           decreasing = TRUE)
i_top5 <- i[1:5]
round(fit_est_change[i_top5, ], 3)
```

We can also specify the parameters to be included. For
example, we may compute the changes and the
`gcd` only for path coefficients.

```{r}
fit_est_change_paths_only <- est_change(fit_rerun,
                                parameters = c("m1 ~ iv1",
                                                "m1 ~ iv2",
                                                "dv ~ m1"))
i <- order(fit_est_change_paths_only[, "gcd"],
           decreasing = TRUE)
round(fit_est_change_paths_only[i[1:5], ], 3)
```

See the help
page of `semfindr::est_change` for further information.

## Raw Changes in Estimates

The standardized changes in parameter may not be easy
to interpret. If the original units are interpretable,
we can compute the *raw* changes, that is, the changes
in parameter estiamtes if a case is deleted, not
standardized by their standard errors. This can be done
by `semfindr::est_change_raw`:

```{r}
fit_est_change_raw <- est_change_raw(fit_rerun)
round(fit_est_change_raw[i_top5, ], 3)
```

If desired, `semfindr::est_change_raw` can also compute the changes
in parameters in the *standardized solution*:

```{r}
fit_est_change_raw_std <- est_change_raw(fit_rerun,
                                         standardized = TRUE)
round(fit_est_change_raw_std[i_top5, ], 3)
```

`semfindr::est_change_raw` also supports computing the changes
for selected parameters:

```{r}
fit_est_change_raw_std_paths_only <- est_change_raw(fit_rerun,
                                          standardized = TRUE,
                                          parameters = c("m1 ~ iv1",
                                                         "m1 ~ iv2",
                                                         "dv ~ m1"))
round(fit_est_change_raw_std_paths_only[i_top5, ], 3)
```

See the help
page of `semfindr::est_change_raw` for further information.

## Mahalanobis Distance

One commonly used measure for identifying outliers is
Mahalanobis distance. `semfindr::mahalanobis_rerun` can be
used to compute the Mahalanobis distance of each case on
all the variables used in the target model:

```{r}
fit_md <- mahalanobis_rerun(fit_rerun)
round(fit_md[i_top5, , drop = FALSE], 3)
```

Note that a case with large Mahalanobis distance is not
necessarily an influential case. Therefore, if time allows,
influence measures should be examined to avoid overlooking
cases that are not extreme but are influential.

## Changes in Fit Measures

Another set of intuitive measures of influence is to compute
goodness of fit mreasures with a case deleted and measure
the changes. This can
be done by `semfindr::fit_measures_change`, which simply
gets any fit measures supported by `lavaan::fitMeasures`
from the results from `semfindr::lavaan_rerun`:

```{r}
fit_mc <- fit_measures_change(fit_rerun,
            fit_measures = c("chisq", "cfi", "tli", "rmsea"))
round(fit_mc[i_top5, ], 3)
```

The value is computed by $M_\textrm{full sample} - M_\textrm{one case removed}$.
Therefore, if the value is positive, the measure is higher
in the full sample. If the value is negative, the measure
is lower in the full sample.

The argument `fit_measures` is passed to `lavaan::fitMeasures`
to sepcify the measures to be computed.

## All-In-One-Function

We can also use `semfindr::influence_stat` to compute the previous
measures. It calls `semfindr::fit_measures_change`, `semfindr::est_change`,
and `semfindr::mahalanobis_rerun` (if requested) and then merges
their results into one matrix:

```{r}
fit_influence <- influence_stat(fit_rerun)
round(fit_influence[i_top5, ], 3)
```

One major use of `semfindr::influence_stat` is to provide information
for the diagnostic plots introduced below.

# Diagnostic Plots

`semfindr` provides several functions to generate diagnostic
plots.

## Generalized Cook's Distance

To visualize the `gcd`, we can plot an index plot with cases
on the horizontal axis and the `gcd` on the vertical axis
using `semfindr:gcd_plot`:

```{r}
gcd_plot(fit_influence, largest_gcd = 3)
```

It accepts an output of `semfindr::influence_stat`.

`largest_gcd` controls the number of cases with the largest
`gcd` to be labelled. The default is 1.

More options of `semfindr::gcd_plot` can be found on its help page.

## Mahalanobis Distance

An index plot can be computed on the
Mahalanobis distance given by `semfindr::influence_stat`:

```{r}
md_plot(fit_influence,
        largest_md = 3)
```

`largest_m` is used to control how many cases with
high Mahalanobis distnace on all the variables in the
fitted model will be labelled. The default is 1.

More options for `md_plot` can be found from its help page.

## Fit Measure vs. Generalized Cook's Distance

To examine how the overall change in parameter estimates
measuresd by `gcd` relates to a selected goodness of fit
measures, `semfindr::gcd_gof_plot` can be used:

```{r}
gcd_gof_plot(fit_influence,
             fit_measure = "rmsea",
             largest_gcd = 3,
             largest_fit_measure = 3)
```

`largest_gcd` determines the number of cases with largest
`gcd` to be labelled, and `largest_fit_measure` determines
the number of cases with largest *absolute* change in
the selected goodness of fit measure to be labelled.
The default is 1 for both arguments.

More options of `semfindr::gcd_gof_plot` can be found from its help
page.

## Bubble Plot

We can also use `semfindr::gcd_gof_md_plot` to plot a bubble plot
of a selected goodness of fit measure against Mahalanobis distance,
with the bubble size determined by generalized Cook's distance:

```{r}
gcd_gof_md_plot(fit_influence,
                fit_measure = "rmsea",
                largest_gcd = 3,
                largest_fit_measure = 3,
                largest_md = 3,
                circle_size = 15)
```

`circle_size` controls the size of the largest bubble.
`largest_gcd`, `largest_fit_measure`, and `largest_md`
controls the number of cases with highest absolute values
one the these measures to be labelled. Their default values
are 1.

More options of `semfindr::gcd_gof_md_plot` can be found from its help
page.