---
title: "Use Case IDs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Use Case IDs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Instead of using row numbers (`case.idx` in the `lavaan` object),
`lavaan_rerun()` from the
package [semfindr](https://sfcheung.github.io/semfindr/)
supports user supplied case IDs.
This can make the output more readable.

```{r}
library(semfindr)
dat <- pa_dat
# Add case id
dat <- cbind(id = paste0("case", seq_len(nrow(dat))), dat)
head(dat)
```

Suppose that the data set has a column of case IDs. A model
is fitted to this data set using `lavaan::sem()`:

```{r}
mod <-
"
m1 ~ iv1 + iv2
dv ~ m1
"
library(lavaan)
fit <- sem(mod, dat)
```

# Rerun *n* Times

We refit the model 100 times, each time with one case
removed. Although the `id` column is not stored in
`lavaan`, it can be supplied through the argument
`case_id`:

```{r echo = FALSE, include = FALSE}
if (file.exists("user_id_fit_rerun.RDS")) {
    fit_rerun <- readRDS("user_id_fit_rerun.RDS")
  } else {
    fit_rerun <- lavaan_rerun(fit, case_id = dat$id)
    saveRDS(fit_rerun, "user_id_fit_rerun.RDS")
  }
```

```{r eval = FALSE}
fit_rerun <- lavaan_rerun(fit, case_id = dat$id)
```

The list of reruns now uses `id` as the names:

```{r}
head(fit_rerun$rerun[1:3])
```

As shown below, most diagnostic functions will use user
supplied case IDs
in their displays, making it easier to locate them in the
original data set.

# Diagnostic Functions

## Standardized Changes in Estimates

```{r}
fit_est_change <- est_change(fit_rerun)
fit_est_change
```
```{r}
fit_est_change_paths_only <- est_change(fit_rerun,
                                parameters = c("m1 ~ iv1",
                                               "m1 ~ iv2",
                                               "dv ~ m1"))
fit_est_change_paths_only
```

## Raw Changes in Estimates

```{r}
fit_est_change_raw <- est_change_raw(fit_rerun)
fit_est_change_raw
```

## Mahalanobis Distance

```{r}
fit_md <- mahalanobis_rerun(fit_rerun)
fit_md
```

## Changes in Fit Measures

```{r}
fit_mc <- fit_measures_change(fit_rerun,
            fit_measures = c("chisq", "cfi", "tli", "rmsea"))
fit_mc
```

## All-In-One-Function

```{r}
fit_influence <- influence_stat(fit_rerun)
fit_influence
```

# Diagnostic Plots

## Generalized Cook's Distance

```{r}
gcd_plot(fit_influence, largest_gcd = 3)
```

## Mahalanobis Distance

```{r}
md_plot(fit_influence,
        largest_md = 3)
```

## Fit Measure vs. Generalized Cook's Distance

```{r}
gcd_gof_plot(fit_influence,
             fit_measure = "rmsea",
             largest_gcd = 3,
             largest_fit_measure = 3)
```

## Bubble Plot

```{r}
gcd_gof_md_plot(fit_influence,
                fit_measure = "rmsea",
                largest_gcd = 3,
                largest_fit_measure = 3,
                largest_md = 3,
                circle_size = 15)
```
