---
title: "Sensitivity Analysis in Multiple-Group Models"
date: "`r Sys.Date()`"
output:
  html_document:
    fig.align: "center"
    toc: true
    number_sections: false
bibliography: references.bib
csl: apa.csl
---

# Goal

This article illustrates how to
identify influential cases in a multiple-group model
using
the [`semfindr`](https://sfcheung.github.io/semfindr/) package
.
This article assumes
that readers have learned how to use the main functions
for single-group models
(See `vignette("semfindr", package = "semfindr")`).

# Dataset and Models

The sample dataset `cfa_dat_mg` in `semfindr` will be used
in this illustration:

```{r}
library(semfindr)
data(cfa_dat_mg)
head(cfa_dat_mg)
table(cfa_dat_mg$gp)
```

The dataset has six variables, `x1` to `x6`. The variable
`gp` denotes the group of a case, with values `"GroupA"`
and `"GroupB"`. Each group has 50 cases and the total
number of cases is 100.

We use assessing measurement invariance as the scenario,
based on the [tutorial](https://lavaan.ugent.be/tutorial/groups.html)
on multiple-group models at the official website of
lavaan.

Three confirmatory factor analytic (CFA) models are to be
fitted. They are all defined by the following model syntax:

```{r}
mod <-
"
f1 =~ x1 + x2 + x3
f2 =~ x4 + x5 + x6
"
```

The first model has no between-group constraints. It tests
configural invariance.

```{r}
library(lavaan)
fit_config <- cfa(mod, cfa_dat_mg,
                  group = "gp")
```

The second model tests weak invariance, with factor loadings
constrained to be equal across groups:

```{r}
fit_weak <- cfa(mod, cfa_dat_mg,
                group = "gp",
                group.equal = "loadings")
```

The third model tests strong invariance, with factor
loadings and item intercepts constrained to be equal
across groups:

```{r}
fit_strong <- cfa(mod, cfa_dat_mg,
                  group = "gp",
                  group.equal = c("loadings", "intercepts"))
```

The three models are compared by `lavaan::lavTestLRT()`:

```{r}
lavTestLRT(fit_config, fit_weak, fit_strong)
lavTestLRT(fit_config, fit_strong)
```

The tests do not reject both strong invariance and
weak invariance.

These are the fit measures of the three models:

```{r}
fm <- c("chisq", "pvalue", "cfi", "tli", "rmsea")
round(data.frame(configural = fitMeasures(fit_config, fm),
                 weak = fitMeasures(fit_weak, fm),
                 strong = fitMeasures(fit_strong, fm)), 3)
```

Although strong invariance is not rejected, the configural
invariance and weak invariance model do not have satisfactory
fit. Interestingly, the most restrictive model, the
strong invariance model, has the best fit. The model chi-square
is not significant.

# Influential Cases

Despite the fit measures, it is still a good practice to
check whether there are any influential cases. Let's start
with the configural invariance model.

## Configural Invariance Model

### Leave-One-Out by `lavaan_rerun()`

The leave-one-out approach is used for illustration on the
configural invariance model, `fit_config`. For this simple
model with only 100 cases, we do not have to use parallel
processing. The model is fitted 100 times, each time with
one case removed:

```{r}
rerun_config <- lavaan_rerun(fit_config)
```

Not shown here, but it is possible that a model may
fail to converge or have inadmissible solutions if removed.
This is especially the case for models with constraints.
The function will report cases leading to nonconvergence
or inadmissible solutions. Influence measures will not be
computed for these cases.

### All-In-One: `influence_stat()`

For preliminary assessment, we can just `influence_stat()`
to compute all commonly used measures of case influence
[@pek_sensitivity_2011],
as well as Mahalanobis distance [@mahalanobis_generalized_1936],
using the output of
`lavaan_rerun()`:

```{r}
inf_config <- influence_stat(rerun_config)
```

The output can then be used for other functions to
examine specific aspects of case influence.

### Mahalanobis Distance

Because the model is a CFA model and all observed variables
are endogenous, it is acceptable to examine Mahalanobis
distance first (see `print.influence_stat()` on options
available when printing the output of `influence_stat()`).

For multiple-group models, Mahalanobis distance for a case
is computed using the mean and covariance matrix of the group
this case belongs to.

```{r}
print(inf_config,
      what = "mahalanobis",
      first = 5)
```

We can visualize the values using `md_plot()`

```{r}
md_plot(inf_config,
        largest_md = 3)
```

Cases 77 and 100 are high on Mahalanobis distance. However,
the differences from other cases are not substantially
large.

### Case Influence on Fit Measures

We then examine case influence on fit measures,
sorted by model chi-squares:

```{r}
print(inf_config,
      what = "fit_measures",
      first = 5,
      sort_fit_measures_by = "chisq")
```

Case 100 has a relatively large influence
on model chi-square.

### Case Influence on Parameter Estimates

Last, we assess casewise influence on parameter estimates
by setting `what` to `"parameters"`. When printed,
by default, cases will be sorted by generalized Cook's
distance [@cook_detection_1977]:

```{r}
print(inf_config,
      what = "parameters",
      first = 5)
```

We can also visualize the generalized Cook's distance
using `gcd_plot()`:

```{r}
gcd_plot(inf_config,
         largest_gcd = 3)
```

The generalized Cook's distance suggests that Case 100
has a strong influence on parameter estimates. It only
affects parameter estimates in `GroupB` because it is
in `GroupB` and the model has no between-group equality
constraints.

We can visualize all three aspects in one plot using
`gcd_gof_md_plot()`:

```{r}
gcd_gof_md_plot(inf_config,
                fit_measure = "chisq",
                circle_size = 15,
                largest_gcd = 3,
                largest_md = 3,
                largest_fit_measure = 3)
```

Case 100 stands out from others when all three
aspects are considered together.

To assess case influence only on factor loadings in
`GroupB` in this model, we can use `est_change()`.
`"=~.GroupB"` denotes all factor loadings in `GroupB`
(see `pars_id()` on how to select parameters):

```{r}
est_change_group <- est_change(rerun_config,
                               parameters = c("=~.GroupB"))
print(est_change_group,
      first = 5)
```

The value of generalized Cook's distance is smaller for Case 100
because only the influence on factor loadings is computed. However,
the difference from those values of other cases is much larger.
This can be visualized using `gcd_plot()`:

```{r}
gcd_plot(est_change_group)
```

### Is Mahalanobis Distance Enough?

Note that the example above illustrates that a case
that is extreme is not necessarily a case that is
influential. Case 100 is influential but it is not the case
with the highest Mahalanobis distance. Admittedly,
Mahalanobis distance does reflect the potential to be
influential. However, in some models, such as models with
regression paths (e.g., a path model), an influential
case may not be an extreme case. Therefore, it is not
enough to examine only Mahalanobis distance.

## Weak Invariance Model and Strong Invariance Model

The above steps can be repeated for the other two models.
They are not shown here for brevity.

# Fit the Models Again Without The Potential Influential Case

Let us fit the three models again, without Case 100.

```{r}
cfa_dat_mg_no100 <- cfa_dat_mg[-100, ]
fit_config_no100 <- cfa(mod, cfa_dat_mg_no100,
                        group = "gp")
fit_weak_no100 <- cfa(mod, cfa_dat_mg_no100,
                      group = "gp",
                      group.equal = "loadings")
fit_strong_no100 <- cfa(mod, cfa_dat_mg_no100,
                        group = "gp",
                        group.equal = c("loadings", "intercepts"))
```
```{r}
lavTestLRT(fit_config_no100, fit_weak_no100, fit_strong_no100)
lavTestLRT(fit_config_no100, fit_strong_no100)
```

Strong invariance is still not rejected if Case 100 is
removed.

```{r}
fm <- c("chisq", "pvalue", "cfi", "tli", "rmsea")
round(data.frame(configural = fitMeasures(fit_config_no100, fm),
                 weak = fitMeasures(fit_weak_no100, fm),
                 strong = fitMeasures(fit_strong_no100, fm)), 3)
```

If Case 100 is removed, all three models fit satisfactorily,
with model chi-squares not significant.

We can check case influence again:

```{r}
rerun_config_no100 <- lavaan_rerun(fit_config_no100)
inf_config_no100 <- influence_stat(rerun_config_no100)
```

For brevity, we only examine the plots:

```{r}
md_plot(inf_config_no100,
        largest_md = 3)
```

```{r}
gcd_plot(inf_config_no100,
         largest_gcd = 3)
```

```{r}
gcd_gof_md_plot(inf_config_no100,
                fit_measure = "chisq",
                circle_size = 15,
                largest_gcd = 3,
                largest_md = 3,
                largest_fit_measure = 3)
```

Although there are still cases with large
generalized Cook's distance, there
are no cases stand out from others,
compared to the analysis with Case 100.

# Final Remarks

Note that the sample dataset above was
artificially created. Case 100 was inserted
to illustrate the workflow and how to use
the functions. In real datasets, it is
usually not that simple in sensitivity
analysis due to the complexity and noise
in real data.

Nevertheless, it is still possible to have
cases easily identifiable. For example,
a case may be influential due to computation
error or reporting error.

Moreover, in real datasets, when one or
more influential cases are identified,
examination of other information is necessary
before deciding what to do with them. We
cannot do this in the example above
because the dataset is artificial. In
real datasets, please see
@aguinis_best-practice_2013.

# Reference